### 2019-05-23 Datelife_paper1.tex
all figures, yay

DateLife palliates this by only using information available from already published
studies, which are ideally constructed using robust information, such as sequence
data and thoughtfully curated fossil calibrations.

Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype \texttt{DateLife} service over a series
of phylotastic hackathons (Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}) at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of \texttt{DateLife}, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
DateLife is the main service for scaling phylogenetic trees in Phylotastic! system
(Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013})
It can be used through an R package , a web interface
(\url{http://www.datelife.org/query/}) and an API.

@@

The basic \texttt{DateLife} workflow is shown in figure \ref{fig:workflow} and consists of:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  A user providing at least two taxon names as input, either as tip labels on a tree, or as a simple comma separated character string. The tree can be in newick or phylo format, and can be with or without branch lengths.
\item
  \texttt{DateLife} then performs a search across its database of peer reviewed and curated chronograms; identifies and gets source trees with at least two matching input names; drops unmatching taxa from positively identified source trees; and finally transforms each source tree to a patristic matrix named by the citation of the original study. This format facilitates and greatly speeds up all further analyses and summarization algorithms.
\item
  The user can obtain different types of summaries from the source data including: a) all source chronograms, b) mrca ages of source chronograms, c) citations of studies where source chronograms were originally published, d) a summary table with all of the above, e) a single summary tree of all source chronograms, and f) a report of succesful matches per input taxon name across source chronograms.
\item
  At this point, users can choose to use all or some source data as calibration points to date a tree of their own making or choosing. 
\item
  Users can also simulate age and/or phylogenetic data of input taxa not found in the database. A variety of algorithms are available for this purpose.
\item
  Finally, users can easily view results graphically as well as construct their own graphs using inbuilt \texttt{DateLife} graphic generators.
\end{enumerate}

\texttt{DateLife}'s chronogram database is currently built from Open Tree of Life (OToL)'s
(Hinchliff et al. \protect\hyperlink{ref-Hinchliff2015}{2015}) tree repository. Among currently existing repositories (e.g., TreeBase,
Dryad), OToL's metadata rich tree store is the only one meeting the requirements
for proper/accurate automatized handling of trees.
Input taxon names accepted by \texttt{DateLife} are binomial species names or clades.
Taxon searches are performed at the species level, so when input names correspond
to higher clades,
\texttt{DateLife} pulls all accepted species names within the
clade from OToL's reference taxonomy to perform the search.
Currently, searches at the infrapsecies level are not allowed, so input names belonging to subspecies or any other infraspecific category are treated
as species.
\texttt{DateLife} also processes input names with the taxon name resolution service (TNRS),
which corrects potentially misspelled names and typos, and standardizes variation
in spelling and synonyms (Boyle et al. \protect\hyperlink{ref-Boyle2013}{2013}), increasing the probability to correctly find the
queried
taxa in \texttt{DateLife}'s chronogram database.

@@

To gather, process, and present information, \texttt{DateLife} builds up from functions
available in several R packages including rotl (Michonneau et al. \protect\hyperlink{ref-Michonneau2016}{2016}), ape (Paradis et al. \protect\hyperlink{ref-Paradis2004}{2004}),
geiger (Harmon et al. \protect\hyperlink{ref-Harmon2008}{2008}), paleotree (Bapst \protect\hyperlink{ref-Bapst2012a}{2012}), bold (Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phytools (Revell \protect\hyperlink{ref-Revell2012}{2012}),
taxize (Chamberlain and Szöcs \protect\hyperlink{ref-Chamberlain2013}{2013}; Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phyloch (Heibl), phylocomr (Ooms and Chamberlain \protect\hyperlink{ref-Ooms2018}{2018})
and rphylotastic (O'Meara et al. \protect\hyperlink{ref-Omeara2019}{2019}).

Details on each step are further developed in \texttt{DateLife}'s R package documentation
\texttt{datelife\ workflow} vignette at (\url{https://LINK}).

@@

\texttt{DateLife}'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor.
We registered variation in computing time relative to number of input names
and \texttt{DateLife} service.
Input processing increases roughly linearly with number of input taxon names, and
increases considerably if tnrs service is activated (Fig. \ref{fig:runtime1}).
Results show that searching time increases linearly with number of input names and
number of chronograms in database.

Summarizing DateLife results processing times

Adding dates processing time

get\_bold\_otol\_tree running time

\texttt{DateLife}'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018}). These tests were run both
locally --using the devtools package (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018})-- and on a public server --via
GitHub-- using the continuous integration tool Travis CI (\textless{}//travis-ci.org\textgreater{}). At
present, unit tests cover more than 50\% (for now) of \texttt{DateLife}'s code (\url{https://codecov.io/gh/phylotastic/datelife}).

@@

In this section we demonstrate the types of outputs that can be obtained with \texttt{datelife}, using the bird family Fringillidae of true finches as example. We performed a higher-taxon search to obtain all data on lineage divergence available from \texttt{datelife}'s database for all recognised species within the Fringillidae (475 spp. according to the Open Tree of Life taxonomy). We found 13 trees across 9 studies (Fig. \ref{fig:schronograms}.

@@

Here, we have shown that DateLife allows
an easy and fast obtention of all publicly available information on taxon ages,
which can be used to generate new data.
This information can be used to account for the effect of phylogenetic signal in
studies of trait evolution; to explore potential speciation and extinction dynamics
of interest within a clade; to obtain a time frame of biogeographical events; for
science communication and outreach, amongst others.
Compared to similar platforms such as time tree of life
and supermart, it offers several advantages.
It is fast;
source data is completely open;
it requires no expert biological knowledge from users for any of its functionalities;
it allows exploration of alternative taxonomic and phylogenetic schemes;
it allows rapid exploration of the effect of alternative divergence time hypothesis;
it allows rapid synthesis in a number of different formats;
it facilitates reproducibility of analyses;

@@

Problems and caveats:
Not many databases, only OToL
Why TreeBase is not very useful for us? Be precise.
Are these chronograms reliable to study evolutionary patterns, such as species diversification?
DateLife can be seen as an open resource to know the current state of knowledge on lineage divergence times.
Whether chronograms obtained using this original data can be used reliably to study complicated patterns of evolution is still uncertain.
If all, la facilidad para obtener hipotesis de tiempo de divergencia nos ayudará a evaluar la capacidad de los cronogramas para estudiar otros fenomenos evolutivos.
Por ahora, no podemos aseverar que estos cronogramas puedan usarse para todo tipo de analisis.

@@

\texttt{Datelife} is free and open source and it can be used through its current website
\url{http://www.datelife.org/query/}, through its R package, and through Phylotastic's project web portal \url{http://phylo.cs.nmsu.edu:3000/}.
\texttt{Datelife}'s website is maintained by RStudio's shiny server and the shiny package open infrastructure, as well as Docker.
\texttt{DateLife}'s R package stable version is available
for installation from the CRAN repository (\url{https://cran.r-project.org/package=datelife})
using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R. Development versions
are available from GitHub repository (\url{https://github.com/phylotastic/datelife})
and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

Code used to generate all versions of this manuscript, the biological examples, as well as the software benchmark can be found in the GitHub repositories \url{https://github.com/LunaSare/datelife_paper1}, \url{https://github.com/LunaSare/datelife_examples}, and \url{https://github.com/LunaSare/datelife_benchmark}, respectively.

@@

We thank colleagues (students and postdocs) at the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for DateLife.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The US National Science Foundation (NSF) for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

@@

\leavevmode\hypertarget{ref-Heibl2008}{}%
Heibl C. PHYLOCH: R language tree plotting tools and interfaces to diverse phylogenetic software packages..

@@

\begin{center}
\textsc{Figure 1}
\end{center}
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.

\begin{center}
\textsc{Figure 2}
\end{center}
Computation time of input processing and search across \texttt{datelife}'s chronogram database.

\begin{center}
\textsc{Figure 3}
\end{center}
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: Barker et al. (\protect\hyperlink{ref-barker2012going}{2012}), 2: Barker et al. (\protect\hyperlink{ref-barker2015new}{2015}), 3: Burns et al. (\protect\hyperlink{ref-burns2014phylogenetics}{2014}), 4: Claramunt and Cracraft (\protect\hyperlink{ref-claramunt2015new}{2015}), 5: Gibb et al. (\protect\hyperlink{ref-gibb2015new}{2015}), 6: Hedges et al. (\protect\hyperlink{ref-Hedges2015}{2015}), 7: Hooper and Price (\protect\hyperlink{ref-hooper2017chromosomal}{2017}), 8: Jetz et al. (\protect\hyperlink{ref-Jetz2012}{2012}), 9: Price et al. (\protect\hyperlink{ref-price2014niche}{2014}).

\begin{center}
\textsc{Figure 4}
\end{center}
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

@@

\begin{figure}[!h]
\includegraphics[width=1\linewidth]{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}

>>>>

\texttt{datelife} palliates this by only using information available from already published
studies, which are ideally constructed using robust information, such as sequence
data and thoughtfully curated fossil calibrations.

Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype \texttt{datelife} service over a series
of phylotastic hackathons (Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}) at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of \texttt{datelife}, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
\texttt{datelife} is the main service for scaling phylogenetic trees in Phylotastic! system
(Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013})
It can be used through an R package , a web interface
(\url{http://www.datelife.org/query/}) and an API.

@@

The basic \texttt{datelife} workflow is shown in figure \ref{fig:workflow} and consists of:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  A user providing at least two taxon names as input, either as tip labels on a tree, or as a simple comma separated character string. The tree can be in newick or phylo format, and can be with or without branch lengths.
\item
  \texttt{datelife} then performs a search across its database of peer reviewed and curated chronograms; identifies and gets source trees with at least two matching input names; drops unmatching taxa from positively identified source trees; and finally transforms each source tree to a patristic matrix named by the citation of the original study. This format facilitates and greatly speeds up all further analyses and summarization algorithms.
\item
  The user can obtain different types of summaries from the source data including: a) all source chronograms, b) mrca ages of source chronograms, c) citations of studies where source chronograms were originally published, d) a summary table with all of the above, e) a single summary tree of all source chronograms, and f) a report of succesful matches per input taxon name across source chronograms.
\item
  At this point, users can choose to use all or some source data as calibration points to date a tree of their own making or choosing. 
\item
  Users can also simulate age and/or phylogenetic data of input taxa not found in the database. A variety of algorithms are available for this purpose.
\item
  Finally, users can easily view results graphically as well as construct their own graphs using inbuilt \texttt{datelife} graphic generators.
\end{enumerate}

\texttt{datelife}'s chronogram database is currently built from Open Tree of Life (OToL)'s
(Hinchliff et al. \protect\hyperlink{ref-Hinchliff2015}{2015}) tree repository. Among currently existing repositories (e.g., TreeBase,
Dryad), OToL's metadata rich tree store is the only one meeting the requirements
for proper/accurate automatized handling of trees.
Input taxon names accepted by \texttt{datelife} are binomial species names or clades.
Taxon searches are performed at the species level, so when input names correspond
to higher clades,
\texttt{datelife} pulls all accepted species names within the
clade from OToL's reference taxonomy to perform the search.
Currently, searches at the infrapsecies level are not allowed, so input names belonging to subspecies or any other infraspecific category are treated
as species.
\texttt{datelife} also processes input names with the taxon name resolution service (TNRS),
which corrects potentially misspelled names and typos, and standardizes variation
in spelling and synonyms (Boyle et al. \protect\hyperlink{ref-Boyle2013}{2013}), increasing the probability to correctly find the
queried taxa in \texttt{datelife}'s chronogram database.

@@

To gather, process, and present information, \texttt{datelife} builds up from functions
available in several R packages including rotl (Michonneau et al. \protect\hyperlink{ref-Michonneau2016}{2016}), ape (Paradis et al. \protect\hyperlink{ref-Paradis2004}{2004}),
geiger (Harmon et al. \protect\hyperlink{ref-Harmon2008}{2008}), paleotree (Bapst \protect\hyperlink{ref-Bapst2012a}{2012}), bold (Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phytools (Revell \protect\hyperlink{ref-Revell2012}{2012}),
taxize (Chamberlain and Szöcs \protect\hyperlink{ref-Chamberlain2013}{2013}; Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phyloch (Heibl \protect\hyperlink{ref-Heibl2008}{2008}), phylocomr (Ooms and Chamberlain \protect\hyperlink{ref-Ooms2018}{2018})
and rphylotastic (O'Meara et al. \protect\hyperlink{ref-Omeara2019}{2019}).

@@

\texttt{datelife}'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor.
We registered variation in computing time of query processing and search through the database relative to number of queried taxon names.
Query processing increases roughly linearly with number of input taxon names, and
increases considerably if TNRS service is activated. Up to ten thousand names can be processed and searched in less than 30 minutes. A name search through the chronogram database with an already processed query can be performed in less than a minute, even with a very large number of taxon names (Fig. \ref{fig:runtime1}).

\texttt{datelife}'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018}). These tests were run both
locally --using the devtools package (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018})-- and on a public server --via
GitHub-- using the continuous integration tool Travis CI (\url{https://travis-ci.org}). At
present, unit tests cover more than 50\% of \texttt{datelife}'s code (\url{https://codecov.io/gh/phylotastic/datelife}).

@@

In this section we demonstrate the types of outputs that can be obtained with \texttt{datelife}, using as an example the bird family Fringillidae of true finches. We performed a higher-taxon search to obtain all data on lineage divergence available from \texttt{datelife}'s database for all recognised species within the Fringillidae (475 spp. according to the Open Tree of Life taxonomy). There are 13 chronograms containing at least two Fringillidae species, published in 9 different studies (Fig. \ref{fig:schronograms1}).
Data from these source chronograms was used to generate two types of summary chronograms, median and SDM. As explained in the \texttt{Description}, data from source chronograms was first summarised into a single distance matrix (using either the median or the SDM method) and then the available node ages were used as calibrations points over a consensus tree topology, to obtain a dated tree with the program BLADJ (Fig. \ref{fig:summaries}). Median summary chronograms are older and have wider variation in maximum ages than chronograms obtained with SDM. In both cases, ages are coherent with source ages.
It is not certain if these chronograms can be used to perform downstream evolutionary analyses. There is currently wide interest in determining this. However, it is certain that these chronograms are useful for\ldots{}

Data from source chronograms was also used to date tree topologies with no branch length information and trees with branch lengths in relative substitution rates (Figs. \ref{fig:cvbladj} and \ref{fig:cvbold}). As a form of cross validation, we used tree topologies from each study and calibrated them using information from all other source chronograms. In the absence of branch length data, the ages of internal nodes were approximately recovered in almost all cases (except for studies 3, and 5; Fig. \ref{fig:cvbladj}). Maximum tree ages were only approximately recovered in one case (study 2; Fig. \ref{fig:cvbladj}).
Branch lengths were successfully generated using the BOLD database for all source chronograms. However, dating with PATHd8 (using congruified calibrations) was only successful in\\
three cases (studies 3, 5, and 9; Fig. \ref{fig:cvbold}). From these, two trees have a different sampling than the original source chronogram, mainly because DNA data for some species is absent from the BOLD. Maximum ages are quite different from source chronograms, but this might be explained also by the differences in sampling between source chronograms and BOLD trees.
More examples and details can be consulted in \url{https://github.com/LunaSare/datelife_examples}.

@@

Here, we have shown that \texttt{datelife} allows
an easy and fast obtention of all publicly available information on taxon ages,
which can be used to generate new data.
This information can be used to account for the effect of phylogenetic signal in
studies of trait evolution; to explore potential speciation and extinction dynamics
of interest within a clade; to obtain a time frame of biogeographical events; for
science communication and outreach, amongst others.
Compared to similar platforms such as time tree of life
and supermart, it offers several advantages.
It is fast;
source data is completely open;
it requires no expert biological knowledge from users for any of its functionalities;
it allows exploration of alternative taxonomic and phylogenetic schemes;
it allows rapid exploration of the effect of alternative divergence time hypothesis;
it allows rapid synthesis in a number of different formats;
it facilitates reproducibility of analyses;

@@

Problems and caveats:
Not many databases, only OToL
Why TreeBase is not very useful for us? Be precise.
Are these chronograms reliable to study evolutionary patterns, such as species diversification?
\texttt{datelife} can be seen as an open resource to know the current state of knowledge on lineage divergence times.
Whether chronograms obtained using this original data can be used reliably to study complicated patterns of evolution is still uncertain.
If all, la facilidad para obtener hipotesis de tiempo de divergencia nos ayudará a evaluar la capacidad de los cronogramas para estudiar otros fenomenos evolutivos.
Por ahora, no podemos aseverar que estos cronogramas puedan usarse para todo tipo de analisis.

@@

\texttt{datelife} is free and open source and it can be used through its current website
\url{http://www.datelife.org/query/}, through its R package, and through Phylotastic's project web portal \url{http://phylo.cs.nmsu.edu:3000/}.
\texttt{datelife}'s website is maintained by RStudio's shiny server and the shiny package open infrastructure, as well as Docker.
\texttt{datelife}'s R package stable version is available
for installation from the CRAN repository (\url{https://cran.r-project.org/package=datelife})
using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R. Development versions
are available from the GitHub repository (\url{https://github.com/phylotastic/datelife})
and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

Code used to generate all versions of this manuscript, the biological examples, as well as the software benchmark can be found in GitHub repositories at \url{https://github.com/LunaSare/datelife_paper1}, \url{https://github.com/LunaSare/datelife_examples}, and \url{https://github.com/LunaSare/datelife_benchmark}, respectively.

@@

We thank colleagues (students and postdocs) at the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for \texttt{datelife}.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The US National Science Foundation (NSF) for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

@@

\leavevmode\hypertarget{ref-Heibl2008}{}%
Heibl C. 2008. PHYLOCH: R language tree plotting tools and interfaces to diverse phylogenetic software packages..

@@

\begin{center}
\textsc{Figure \ref{fig:workflow}}
\end{center}
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with \texttt{datelife}, via the R package or through the website. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.

\begin{center}
\textsc{Figure \ref{fig:runtime1}}
\end{center}
Computation time of query processing and search across \texttt{datelife}'s chronogram database relative to number of input taxon names. We sampled N names from the class Aves for each cohort 100 times and then performed a search with query processing not using the Taxon Names Resoultion Service (TNRS; dark gray), and using TNRS (light gray). We also performed a search using the already processed query for comparison (light blue).

\begin{center}
\textsc{Figure \ref{fig:schronograms1}}
\end{center}
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: Barker et al. (\protect\hyperlink{ref-barker2012going}{2012}), 2: Barker et al. (\protect\hyperlink{ref-barker2015new}{2015}), 3: Burns et al. (\protect\hyperlink{ref-burns2014phylogenetics}{2014}), 4: Claramunt and Cracraft (\protect\hyperlink{ref-claramunt2015new}{2015}), 5: Gibb et al. (\protect\hyperlink{ref-gibb2015new}{2015}), 6: Hedges et al. (\protect\hyperlink{ref-Hedges2015}{2015}), 7: Hooper and Price (\protect\hyperlink{ref-hooper2017chromosomal}{2017}), 8: Jetz et al. (\protect\hyperlink{ref-Jetz2012}{2012}), 9: Price et al. (\protect\hyperlink{ref-price2014niche}{2014}).

\begin{center}
\textsc{Figure \ref{fig:summaries}}
\end{center}
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarising information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\begin{center}
\textsc{Figure \ref{fig:cvbladj}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees without branch lengths dated using BLADJ. The dating analysis can only be performed in trees with more than 2 tips, thus excluding chronogram from study 4; its data was still used as calibration for the other source chronograms.

\begin{center}
\textsc{Figure \ref{fig:cvbold}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees with branch length reconstructed with data from the Barcode of Life Database (BOLD) dated using PATHd8. We could construct a tree with branch lengths for all source chronograms. However, dating with PATHd8 was only successfull in three source chronograms shown here.

@@

\begin{figure}[!h]
\includegraphics[width=1\linewidth]{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms1}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics{fig_summaries.pdf}
%\includegraphics{fig_summaries.pdf}
\caption{}
\label{fig:summaries}
\end{figure}

\begin{figure}[!h]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}

\begin{figure}[!h]
\includegraphics{fig_crossval_boldsumm.pdf}
\caption{}
\label{fig:cvbold}
\end{figure}

### 2019-05-23 Datelife_paper1_figures_noyaml.tex
all figures, yay

\leavevmode\hypertarget{ref-barker2012going}{}%
Barker F.K., Burns K.J., Klicka J., Lanyon S.M., Lovette I.J. 2012. Going to extremes: Contrasting rates of diversification in a recent radiation of new world passerine birds. Systematic biology. 62:298--320.

\leavevmode\hypertarget{ref-barker2015new}{}%
Barker F.K., Burns K.J., Klicka J., Lanyon S.M., Lovette I.J. 2015. New insights into new world biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk: Ornithological Advances. 132:333--348.

\leavevmode\hypertarget{ref-burns2014phylogenetics}{}%
Burns K.J., Shultz A.J., Title P.O., Mason N.A., Barker F.K., Klicka J., Lanyon S.M., Lovette I.J. 2014. Phylogenetics and diversification of tanagers (passeriformes: Thraupidae), the largest radiation of neotropical songbirds. Molecular Phylogenetics and Evolution. 75:41--77.

\leavevmode\hypertarget{ref-claramunt2015new}{}%
Claramunt S., Cracraft J. 2015. A new time tree reveals earth history's imprint on the evolution of modern birds. Science advances. 1:e1501005.

\leavevmode\hypertarget{ref-gibb2015new}{}%
Gibb G.C., England R., Hartig G., McLenachan P.A., Taylor Smith B.L., McComish B.J., Cooper A., Penny D. 2015. New zealand passerines help clarify the diversification of major songbird lineages during the oligocene. Genome biology and evolution. 7:2983--2995.

\leavevmode\hypertarget{ref-Hedges2015}{}%
Hedges S.B., Marin J., Suleski M., Paymer M., Kumar S. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution. 32:835--845.

\leavevmode\hypertarget{ref-hooper2017chromosomal}{}%
Hooper D.M., Price T.D. 2017. Chromosomal inversion differences correlate with range overlap in passerine birds. Nature ecology \& evolution. 1:1526.

\leavevmode\hypertarget{ref-Jetz2012}{}%
Jetz W., Thomas G., Joy J.J., Hartmann K., Mooers A. 2012. The global diversity of birds in space and time. Nature. 491:444--448.

\leavevmode\hypertarget{ref-price2014niche}{}%
Price T.D., Hooper D.M., Buchanan C.D., Johansson U.S., Tietze D.T., Alström P., Olsson U., Ghosh-Harihar M., Ishtiaq F., Gupta S.K., others. 2014. Niche filling slows the diversification of himalayan songbirds. Nature. 509:222.

@@

\begin{center}
\textsc{Figure 1}
\end{center}
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.

\begin{center}
\textsc{Figure 2}
\end{center}
Computation time of input processing and search across \texttt{datelife}'s chronogram database.

\begin{center}
\textsc{Figure 3}
\end{center}
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: Barker et al. (\protect\hyperlink{ref-barker2012going}{2012}), 2: Barker et al. (\protect\hyperlink{ref-barker2015new}{2015}), 3: Burns et al. (\protect\hyperlink{ref-burns2014phylogenetics}{2014}), 4: Claramunt and Cracraft (\protect\hyperlink{ref-claramunt2015new}{2015}), 5: Gibb et al. (\protect\hyperlink{ref-gibb2015new}{2015}), 6: Hedges et al. (\protect\hyperlink{ref-Hedges2015}{2015}), 7: Hooper and Price (\protect\hyperlink{ref-hooper2017chromosomal}{2017}), 8: Jetz et al. (\protect\hyperlink{ref-Jetz2012}{2012}), 9: Price et al. (\protect\hyperlink{ref-price2014niche}{2014}).

\begin{center}
\textsc{Figure 4}
\end{center}
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

>>>>

@@

\begin{center}
\textsc{Figure \ref{fig:workflow}}
\end{center}
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with \texttt{datelife}, via the R package or through the website. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.

\begin{center}
\textsc{Figure \ref{fig:runtime1}}
\end{center}
Computation time of query processing and search across \texttt{datelife}'s chronogram database relative to number of input taxon names. We sampled N names from the class Aves for each cohort 100 times and then performed a search with query processing not using the Taxon Names Resoultion Service (TNRS; dark gray), and using TNRS (light gray). We also performed a search using the already processed query for comparison (light blue).

\begin{center}
\textsc{Figure \ref{fig:schronograms1}}
\end{center}
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: Barker et al. (\protect\hyperlink{ref-barker2012going}{2012}), 2: Barker et al. (\protect\hyperlink{ref-barker2015new}{2015}), 3: Burns et al. (\protect\hyperlink{ref-burns2014phylogenetics}{2014}), 4: Claramunt and Cracraft (\protect\hyperlink{ref-claramunt2015new}{2015}), 5: Gibb et al. (\protect\hyperlink{ref-gibb2015new}{2015}), 6: Hedges et al. (\protect\hyperlink{ref-Hedges2015}{2015}), 7: Hooper and Price (\protect\hyperlink{ref-hooper2017chromosomal}{2017}), 8: Jetz et al. (\protect\hyperlink{ref-Jetz2012}{2012}), 9: Price et al. (\protect\hyperlink{ref-price2014niche}{2014}).

\begin{center}
\textsc{Figure \ref{fig:summaries}}
\end{center}
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarising information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\begin{center}
\textsc{Figure \ref{fig:cvbladj}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees without branch lengths dated using BLADJ. The dating analysis can only be performed in trees with more than 2 tips, thus excluding chronogram from study 4; its data was still used as calibration for the other source chronograms.

\begin{center}
\textsc{Figure \ref{fig:cvbold}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees with branch length reconstructed with data from the Barcode of Life Database (BOLD) dated using PATHd8. We could construct a tree with branch lengths for all source chronograms. However, dating with PATHd8 was only successful in three source chronograms shown here.
